---
title: "Worksheet-4c in R"
author: "Necel Kate Noblezada"
date: "2024-10-30"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Use the dataset mpg

# 1.
# a.
```{r}
mpg_data <- read.csv("mpg.csv")

```
# b.
```{r}
str(mpg_data)
```
The categorical variables are manufacturer, model, trans, drv, fl, and class.

# c.
The continous variables are displ, year, cyl, cty, and hwy.

# 2.
```{r}
library(dplyr)
library(ggplot2)

manufacturer_count <- mpg %>%
  group_by(manufacturer) %>%
  summarise(num_models = n_distinct(model)) %>%
  arrange(desc(num_models))

manufacturer_count[1, ]

model_count <- mpg %>%
  group_by(model) %>%
  summarise(num_variations = n()) %>%
  arrange(desc(num_variations))

model_count[1, ]
```

# a.
```{r}
unique_models <- mpg %>%
  group_by(manufacturer) %>%
  summarise(unique_models_count = n_distinct(model))

unique_models
```
# b.
```{r}
barplot(
  unique_models$unique_models_count,
  names.arg = unique_models$manufacturer,
  las = 2,                    # Make x-axis labels vertical for better readability
  col = "lightblue",
  main = "Number of Unique Models by Manufacturer",
  xlab = "Manufacturer",
  ylab = "Unique Models Count"
)

ggplot(unique_models, aes(x = reorder(manufacturer, -unique_models_count), y = unique_models_count)) +
  geom_bar(stat = "identity", fill = "blue") +
  labs(title = "Number of Unique Models by Manufacturer", x = "Manufacturer", y = "Unique Models Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
# c. Which are continuous variables?
- displ
- year
- cyt
- hwy

# 2. Which manufacturer has the most models in this data set? Which model has the most variations?
 Show your answer
```{r}
 library(dplyr)
```
```{r}
 manufacturer_model_count <- mpg %>%
 group_by(manufacturer) %>%
 summarise(model_count = n_distinct(model)) %>%
 arrange(desc(model_count))
 most_models <- manufacturer_model_count[1, ]
 model_variation_count <- mpg %>%
 group_by(model) %>%
 summarise(variation_count = n()) %>%
 arrange(desc(variation_count))
 most_variations_model <- model_variation_count[1, ]
 most_models
```
# a. Group the manufacturers and find the unique models. Show your codes and result
```{r}
unique_models <- mpg %>%
 group_by(manufacturer) %>%
 summarise(unique_models = list(unique(model))) %>%
 arrange(manufacturer)
 print(unique_models)
```
# b. Graph the result by using plot() and ggplot(). Write the codes and its result.
```{r}
 model <- mpg %>%
 group_by(manufacturer) %>%
 summarise(unique_count = n_distinct(model)) %>%
 arrange(desc(unique_count))
 barplot(model$unique_count,
 names.arg = model$manufacturer,
 las = 2,
 col = "lightgrey",
 main = "Number of Unique Models by Manufacturer",
 xlab = "Manufacturer",
 ylab = "Number of Unique Models")
```
```{r}
 library(ggplot2)

ggplot(model, aes(x = reorder(manufacturer, unique_count), y = unique_count)) +
  geom_bar(stat = "identity", fill = "lightgrey") +
  coord_flip() +
  labs(title = "Number of Unique Models by Manufacturer",
       x = "Manufacturer",
       y = "Number of Unique Models") +
  theme_minimal()

```

# 2. Same dataset will be used. You are going to show the relationship of the model and the manufacturer.
# a. What does ggplot(mpg, aes(model, manufacturer))+geom_point() show?
# 
# b. For you, is it useful? If not,how could you modify the data to make it more informative? Usefulness:
# 

# 3. Plot the model and the year using ggplot(). Use only the top 20 observations. Write the codes and its results.
```{r}
 Top_20<-mpg%>%
 arrange(desc(cty)) %>%
 head(20)
 ggplot(Top_20, aes(x=model,y=year))+
 geom_point() +
 labs(title="Top20ModelsbyYear",x="Model",y="Year") +
 theme(axis.text.x=element_text(angle=90,hjust=0.5))
```

# 4. 
```{r}
car_counts <- mpg %>%
  group_by(model) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
```
# a.
```{r}
top_20 <- head(car_counts, 20)

ggplot(top_20, aes(x = reorder(model, count), y = count, fill = count)) +
  geom_bar(stat = "identity") +
  labs(title = "Top 20 Car Models by Count",
       x = "Car Model",
       y = "Number of Cars") +
  scale_fill_gradient(low = "green", high = "red") + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


# b.
```{r}
ggplot(top_20, aes(x = reorder(model, count), y = count, fill = count)) +
  geom_bar(stat = "identity") +
  labs(title = "Top 20 Car Models by Count (Horizontal)",
       x = "Number of Cars",
       y = "Car Model") +
  scale_fill_gradient(low = "green", high = "red") +  # Color gradient
  theme_minimal() +
  coord_flip()
```

# 5.
```{r}
ggplot(mpg, aes(x = cyl, y = displ, color = displ)) +
  geom_point(size = 3, alpha = 0.7) +  # Adjust point size and transparency
  labs(title = "Relationship between No. of Cylinders and Engine Displacement",
       x = "Number of Cylinders",
       y = "Engine Displacement (L)") +
  scale_color_gradient(low = "green", high = "red") + 
  theme_minimal()
```
# a. 
The more number of cylinders the higher the displacement. It shows a positive relationship.

# 6. 
```{r}
ggplot(mpg, aes(x = displ, y = hwy, color = cty)) +
  geom_point(size = 3, alpha = 0.7) + 
  labs(title = "Relationship between Engine Displacement and Highway MPG",
       x = "Engine Displacement (L)",
       y = "Highway Miles per Gallon (MPG)") +
  scale_color_gradient(low = "green", high = "red") +
  theme_minimal()
```

# 6. 
```{r}
traffic_data <- read.csv("traffic.csv")

```
# a. 
```{r}
length(traffic_data)

variable_names <- names(traffic_data)

variable_names
```
# b. 
```{r}
unique_junctions <- unique(traffic_data$Junction)

junctions_dataframes <- list()

for (junction in unique_junctions) {
  junctions_dataframes[[junction]] <- traffic_data %>%
    filter(Junction == junction)
}

head(junctions_dataframes[[1]])
```

# c. 
```{r}
ggplot(traffic_data, aes(x = DateTime, y = Vehicles, color = Junction)) +
  geom_line() +
  labs(title = "Traffic Count by Junction",
       x = "Date and Time",
       y = "Number of Vehicles") +
  theme_minimal() +
  theme(legend.position = "right") 
```
# 7. 
```{r}
library(readxl)

alexa_data <- read_excel("alexa_file.xlsx")
```
# a. 
```{r}
num_observations <- nrow(alexa_data)

num_columns <- ncol(alexa_data)

num_observations
num_columns
```
# b. 
```{r}
alexa_data$rating <- as.numeric(as.character(alexa_data$rating))
alexa_data$verified_reviews <- as.numeric(as.character(alexa_data$verified_reviews))

sum(is.na(alexa_data$rating))  
sum(is.na(alexa_data$verified_reviews)) 

variation_totals <- alexa_data %>%
  group_by(variation) %>%  
  summarize(Total_Rating = sum(rating, na.rm = TRUE),  
            Total_Verified_Reviews = sum(verified_reviews, na.rm = TRUE))

print(variation_totals)
```
# c. 
```{r}
ggplot(variation_totals, aes(x = variation, y = Total_Rating)) +
  geom_bar(stat = "identity", fill = "orange") +
  labs(title = "Total Rating by Variation",
       x = "Variation",
       y = "Total Rating") +
  theme_minimal()
```
The graph shows that each bar corresponds to the total rating for a specific variation. Although the names on the x axis overlaps.
# d. 
```{r}
reviews_by_date <- alexa_data %>%
  group_by(date) %>%  # Group by date
  summarize(Total_Verified_Reviews = n())

ggplot(reviews_by_date, aes(x = date, y = Total_Verified_Reviews)) +
  geom_line(color = "blue") +  
  labs(title = "Total Verified Reviews Over Time", x = "Date", y = "Total Verified Reviews") +
  theme_minimal()  
```

# e. 
```{r}
average_ratings <- alexa_data %>%
  group_by(variation) %>%  
  summarize(Average_Rating = mean(rating, na.rm = TRUE)) 

highest_rating_variation <- average_ratings %>%
  filter(Average_Rating == max(Average_Rating)) 

ggplot(average_ratings, aes(x = reorder(variation, -Average_Rating), y = Average_Rating)) +
  geom_bar(stat = "identity", fill = "lightblue") +  
  labs(title = "Average Ratings by Variation",
       x = "Variation",
       y = "Average Rating") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```

