---
title: "Worksheet-4c in R"
author: "Necel Kate Noblezada"
date: "2024-10-30"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Use the dataset mpg
# A data frame with 234 rows and 11 variables:
#’ \describe{
#’ \item{manufacturer}{manufacturer name}
#’ \item{model}{model name}
#’ \item{displ}{engine displacement, in litres}
#’ \item{year}{year of manufacture}
#’ \item{cyl}{number of cylinders}
#’ \item{trans}{type of transmission}
#’ \item{drv}{the type of drive train, where f = front-wheel drive, r = rear wheel drive, 4 = 4wd}
#’ \item{cty}{city miles per gallon}
#’ \item{hwy}{highway miles per gallon}
#’ \item{fl}{fuel type}
#’ \item{class}{"type" of car}
#’ }
# "mpg"
# - Download and open the mpg file. Upload it to your OWN environment
# a. Show your solutions on how to import a csv file into the environment.
```{r}
library(readr)
 mpg<-read_csv("mpg.csv")
```
# b. Which variables from mpg dataset are categorical?
# manufacturer-The carmanufacturer (e.g.,Audi, Chevrolet).-model- trans- drv- cyl- fl- class
manufacturer
# c. Which are continuous variables?
- displ
- year
- cyt
- hwy

# 2. Which manufacturer has the most models in this data set? Which model has the most variations?
 Show your answer
```{r}
 library(dplyr)
```
```{r}
 manufacturer_model_count <- mpg %>%
 group_by(manufacturer) %>%
 summarise(model_count = n_distinct(model)) %>%
 arrange(desc(model_count))
 most_models <- manufacturer_model_count[1, ]
 model_variation_count <- mpg %>%
 group_by(model) %>%
 summarise(variation_count = n()) %>%
 arrange(desc(variation_count))
 most_variations_model <- model_variation_count[1, ]
 most_models
```
# a. Group the manufacturers and find the unique models. Show your codes and result
```{r}
unique_models <- mpg %>%
 group_by(manufacturer) %>%
 summarise(unique_models = list(unique(model))) %>%
 arrange(manufacturer)
 print(unique_models)
```
# b. Graph the result by using plot() and ggplot(). Write the codes and its result.
```{r}
 model <- mpg %>%
 group_by(manufacturer) %>%
 summarise(unique_count = n_distinct(model)) %>%
 arrange(desc(unique_count))
 barplot(model$unique_count,
 names.arg = model$manufacturer,
 las = 2,
 col = "lightgrey",
 main = "Number of Unique Models by Manufacturer",
 xlab = "Manufacturer",
 ylab = "Number of Unique Models")
```
```{r}
 ggplot(model, aes(x = reorder(manufacturer, unique_count), y = unique_count)) +
 geom_bar(stat = "identity", fill = "lightgrey") +
 coord_flip() +
 labs(title = "Number of Unique Models by Manufacturer",
 x = "Manufacturer",
 y = "Number of Unique Models") +
 theme_minimal()
```

# 2. Same dataset will be used. You are going to show the relationship of the model and the manufacturer.
# a. What does ggplot(mpg, aes(model, manufacturer))+geom_point() show?
# 
# b. For you, is it useful? If not,how could you modify the data to make it more informative? Usefulness:
# 

# 3. Plot the model and the year using ggplot(). Use only the top 20 observations. Write the codes and its results.
```{r}
 Top_20<-mpg%>%
 arrange(desc(cty)) %>%
 head(20)
 ggplot(Top_20, aes(x=model,y=year))+
 geom_point() +
 labs(title="Top20ModelsbyYear",x="Model",y="Year") +
 theme(axis.text.x=element_text(angle=90,hjust=0.5))
```



