---
title: "Worksheet-1in R"
author: "Necel Kate Noblezada"
date: "2024-09-25"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Create a vector using : operator
# a. Sequence from -5 to 5. Write the R code and its output. Describe its output.
```{r}
number <- -5:5
print(number)
```
# The output is sequenced from -5 to 5.
# b. x <- 1:7. What will be the value of x?
```{r}
x <- 1:7
print(x)
```
# The value of x will be 1 to 7.

# 2. Create a vector using seq() function
# seq(1, 3, by=0.2) # specify step size
# Write the R script and its output. Describe the output.
```{r}
number <- seq(1, 3, by = 0.2)
print(number)
```
# The output is a vector of numbers starting from 1 and ending at 3.

# 3. A factory has a census of its workers. There are 50 workers in total. The following
list shows their ages: 34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27,
22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 43, 53, 41, 51, 35,
24,33, 41, 53, 40, 18, 44, 38, 41, 48, 27, 39, 19, 30, 61, 54, 58, 26,
18.
```{r}
ages <- c(34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27,
22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 43, 53, 41, 51, 35,
24,33, 41, 53, 40, 18, 44, 38, 41, 48, 27, 39, 19, 30, 61, 54, 58, 26,
18)
```
# a. Access 3rd element, what is the value?
```{r}
third <- ages[3]
print(third)
```
# The value of the third element is 22.
# b. Access 2nd and 4th element, what are the values?
```{r}
second_fourth <- ages[c(2,4)]
print(second_fourth)
```
# The value for second is 28 and the value for the fourth element is 36
# c. Access all but the 4th and 12th element is notincluded. Write the R script and its output.
```{r}
all <- ages[-c(4,12)]
print(all)
```

# 4. Create a vector x <- c("first"=3, "second"=0, "third"=9). Then named the vector, names(x).
```{r}
x <- c("first" = 3, "second" = 0, "third" = 9)
print(x)
```
# a. Print the results. Then access x[c("first", "third")]. Describe the output.
```{r}
names <- x[c("first", "third")]
print(names)
```
# The outputs are the first and third elements.

# 5. Create a sequence x from -3:2.
# a. Modify 2nd element and change it to 0;
# x[2] <- 0
# x
# Describe the output.
```{r}
x <- -3:2
x[2] <- 0
print(x)
```
# The output vector is `-3, 0, -1, 0, 1, 2` after replacing the second element with `0`.

# 6. The following data shows the diesel fuel purchased by Mr. Cruz.
# a. Create a data frame for month, price per liter (php) and purchase-quantity (liter). Write the R scripts and its output.
```{r}
month <- c("Jan", "Feb","March", "April", "May", "June")
php <- c(52.52, 57.25, 60.00, 65.00, 74.25, 54.00)
liter <- c(25, 30, 40, 50, 10, 45)

fuel_data <- data.frame(Month = month, Php = php, Liter = liter)
print(fuel_data)
```
# b. What is the average fuel expenditure of Mr. Cruz from Jan to June? Note: Use ‘weighted.mean(liter, purchase)‘. Write the R scripts and its output.
```{r}
average_exp <- weighted.mean(php, liter)
print(average_exp)
```
# The average fuel expenditure of Mr. Cruz from January to June is 59.00 PHP per liter.

# 7. R has actually lots of built-in datasets. For example, the rivers data “gives the lengths (in miles) of 141 “major” rivers in North America, as compiled by the US Geological Survey”.
# a. Type “rivers” in your R console.
# Create a vector data with 7 elements, containing the number of elements (length) in rivers, their sum (sum), mean (mean), median(median), variance(var), standard deviation(sd), minimum (min) and maximum (max).
#data <- c(length(rivers), sum(rivers), mean(rivers), median(rivers), var(rivers),sd(rivers), min(rivers), max(rivers))
```{r}
data <- c(length(rivers), sum(rivers), mean(rivers), median(rivers), var(rivers), var(rivers), sd(rivers), min(rivers), max(rivers))
print(data)
```
# b. What are the results?
# The results show that there are 141 rivers with a total length of 34,034 miles, an average length of 241.20 miles, a median of 200 miles, variance of 7861.65, standard deviation of 88.75 miles, a minimum length of 12 miles, and a maximum length of 1,248 miles.

# 8. The table below gives the 25 most powerful celebrities and their annual pay as ranked by the editions of Forbes magazine and as listed on the Forbes.com website.
# a. Create vectors according to the above table. Write the R scripts and its output.
```{r}
celebrity <- c("Tom Cruise", "Rolling Stones", "Oprah Winfrey", "U2", "Tiger Woods", "Steven Spielberg", "Howard Stern", "50 Cent", "Cast of the Sopranos", "Dan Brown", " Bruce Springsteen", "Donald Trump", "Muhammad Ali", "Paul McCartney", "George Lucas", "Elton John", "David Letterman", "Phil Mickelson", "J.K Rowling", "Brad Pitt", "Peter Jackson", "Dr. Phil McGraw", "Jay Lenon", "Celine Dion", "Kobe Bryant")
pay <- c(67, 90, 225, 110, 90, 332, 302, 41, 52, 88, 55, 44, 55, 40, 233, 34, 40, 47, 75, 25, 39, 45, 32, 40, 31)
power_ranking <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25)
celebrity_data <- data.frame(Celebrity = celebrity, Pay = pay, Power_ranking = power_ranking)
print(celebrity_data)
```
# b. Modify the power ranking and pay of J.K. Rowling. Change power ranking to 15 and pay to 90. Write the R scripts and its output.
```{r}
power_ranking[19] <- 15 
pay[19] <- 90
celebrity_data <- data.frame(Celebrity = celebrity, Pay = pay, Power_ranking = power_ranking)
print(celebrity_data)
```
# c. Create an excel file from the table above and save it
as csv file(PowerRanking). Import the csv file into
the RStudio. What is the R script?
```{R}
write.csv(celebrity_data, file = "PowerRanking.csv", row.names = FALSE)
celebrity_data_imported <- read.csv("PowerRanking.csv")
celebrity_data_imported
``` 
# d. Access the rows 10 to 20 and save it as Ranks.RData.
Write the R script and its output.
```{r}
ranks <- celebrity_data[10:20,]
save(ranks, file = "Ranks.RData")
print(ranks)
```
# e. Describe its output.
# The output will display a table of rows 10 to 20 from the `celebrity_data_imported` data frame, showing columns like Name, Rank, Score, and Category for each celebrity in that range.

# 9. Download the Hotels-Vienna https://tinyurl.com/Hotels-Vienna
# a. Import the excel file into your RStudio.
# What is the R script?
```{r}
Hotels_Vienna_data <- read.csv("hotels-vienna.xlsx - hotels-vienna.csv")
```
# b. How many dimensions does the dataset have? What is the R script? What is its output?
```{r}
dim(Hotels_Vienna_data)
```
# c. Select columns country, neighbourhood, price, stars, accomodation_type, and ratings. Write the R script.
```{r}
library(dplyr)
selected_columns <- Hotels_Vienna_data %>% select(country, neighbourhood, price, stars, accommodation_type, rating)
selected_columns
```
# d. Save the data as **new.RData to your RStudio. Write the
R script.
```{r}
save(selected_columns, file = "new.RData")
```
# e. Display the first six rows and last six rows of the new.RData. What is the R script?
```{r}
load("new.RData")

first_six_rows <- head(selected_columns)
print(first_six_rows)

last_six_rows <- tail(selected_columns)
print(last_six_rows)
```

# 10. Create a list of ten (10) vegetables you ate during your lifetime. If none, just list down.
# a. Write the R scripts and its output.
```{r}
vegetables <- c("zucchini", "okra", "cauliflower", "potato", "cabbage", "carrot", "lettuce", "broccoli", "mushroom", "eggplant")
print(vegetables)
```
# b. Add 2 additional vegetables after the last vegetables in the list. What is the R script and its output?
```{r}
vegetables <- c(vegetables, "garlic","radish")
print(vegetables)
```
# c. Add 4 additional vegetables after index 5. How manydatapoints does your vegetable list have? What is the R script and its output?
```{r}
vegetables <- append(vegetables, c("beans", "lemongrass", "leek", "cilantro"), after = 5)
print(vegetables)
length(vegetables)
```
# d. Remove the vegetables in index 5, 10, and 15. How many vegetables were left? Write the codes and its output.
```{r}
vegetables <- vegetables[-c(5, 10, 15)]
print(vegetables)
length(vegetables)
```

