---
title: "Worksheet-1 in R"
author: "Necel Kate Noblezada"
date: "2024-09-04"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
# 1. Set up a vector named age
```{r}
age <- c(34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27, 22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 42, 53, 41, 51, 35, 24, 33, 41)
```
# a. How many data points?
# answer: 34 data points
```{r}
length(age)
```


# 2. Find the reciprocal of the values for age.
```{r}
reciprocal<-1/age
reciprocal
```

# 3. Assign also new_age <- c(age, 0, age).
```{r}
new_age <- c(age, 0, age)
new_age
```
# What happen to the new_age?
# answer: It duplicates the age vector with 0 in the middle.


# 4. Sort the values for age.
```{r}
sort(age)
```

# 5. Find the minimum and maximum value for age.
```{r}
min(age)
max(age)
```


# 6. Set up a vector named data
```{r}
data<-c(2.4, 2.8, 2.1, 2.5, 2.4, 2.2, 2.5, 2.3, 2.5, 2.3, 2.4, 2.7)
```
# a. How many data points?
# answer: 12 data points
```{r}
length(data)
```


# 7. Generate a new vector for data where you double every value of the data.
```{r}
new_vector <- data
data * 2
```
# What happened to the data?
# answer: The data vector results in a new vector where each element is twice the original value.


# 8. Generate a sequence for the following scenario:
# 8.1 - Integers from 1 to 100.
```{r}
a <- seq(1:100)
```
8.2 - Numbers from 20 to 60.
```{r}
l <- seq(20,60)
```
8.3 - Mean of numbers from 20 to 60.
```{r}
i <- mean(20,60)
```
8.4 - Sum of numbers from 51 to 91.
```{r}
k <- sum(51:91)
```
8.5 - Integers from 1 to 1,000.
```{r}
t <- seq(1:1000)
```
# a. How many data points from 8.1 to 8.4?
# answer: 143 data points
```{r}
length1 <- length(a)
length2 <- length(l)
length3 <- length(i)
length4 <- length(k)
sum (length1+length2+length3+length4)
```
# b. Write the R code and its output from 8.1 to 8.4.
```{r}
a <- seq(1:100)
l <- seq(20,60)
i <- mean(20,60)
k <- sum(51:91)
```    
# c. For 8.5 find only maximum data points until 10.
```{r}
t <- 1:1000
answer <- max(t[t <- 10])
answer
```


# 9. *Print a vector with the integers between 1 and 100 that are not divisible by 3, 5 and 7 using filter option.
```{r}
Filter(function(i) { all(i %% c(3,5,7) != 0) }, seq(100))
```


# 10. Generate a sequence backwards of the integers from 1 to 100.
```{r}
rev(seq(1:100))
```


# 11. List all the natural numbers below 25 that are multiples of 3 or 5.
```{r}
upper_limit <- 25
num <- 1:(upper_limit - 1)
multiples <- num[num %% 3 == 0 | num %% 5 == 0]
multiples
```
# a. How many data points from 10 to 11?
# answer: 11 data points
```{r}
length(multiples)
```
# b. Write the R code and its output from 10 and 11.
```{r}
dp10 <- length(multiples)
dp11 <- 1

total <- dp10 + dp11
total
```


# 12. 
```{r}
x <- {0 + x + 5 + 3} 
```

# Describe the output.
# answer: he line is incomplete. Thus, it caused an error. 


# 13. Set up a vector named score
```{r}
score <- c(72, 86, 92, 63, 88, 89, 91, 92, 75,
75, 77)

x2 <- score[2]
x3 <- score[3]

print(x2)
print(x3)
```


# 14. Create a vector
```{r}
a <- c(1,2,NA,4,NA,6,7)

print(a, na.print = "-999")
```
# b. describe the output.
# answer: The output displays the vector a with NA values shown as -999, while the actual vector remains unchanged with NA values.



# 15. 
```{r}
name = readline(prompt="Input your name: ")
age = readline(prompt="Input your age: ")
print(paste("My name is",name, "and I am",age ,"years old."))
print(R.version.string)
```
# What is the output of the above code?
# answer: The output of the code is a message stating the user's name and age followed by the R version string, e.g., `"My name is "name" and I am "age" years old."` and `"R version 4.3.1 (2024-06-01)"`.

