---
title: "RWorksheet_noblezada#1"
author: "Necel Kate Noblezada"
date: "2024-09-04"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
# 1.
```{r}
age=c(34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27, 22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 42, 53, 41, 51, 35, 24, 33, 41)
```

#a. How many data points?
# 34 data points
```{r}
length(age)
```
#2.
```{r}
reciprocal<-1/age
reciprocal
```
#3. 
#What happen to the new_age?
# 
```{r}
new_age <- c(age, 0, age)
new_age
```
#4.
```{r}
sort(age)
```
#5.
```{r}
min(age)
max(age)
```

#6.
```{r}
data<-c(2.4, 2.8, 2.1, 2.5, 2.4, 2.2, 2.5, 2.3, 2.5, 2.3, 2.4, 2.7)
```
#a. How many data points?
# 12 data points
```{r}
length(data)
```
#7.

```{r}
new_vector <- data
data * 2
```
#What happen to the data?
# the data vector results in a new vector where each element is twice the original value.

8.1
```{r}
a <- seq(1:100)
```
8.2
```{r}
l <- seq(20,60)
```

8.3
```{r}
i <- mean(20,60)
```

8.4
```{r}
k <- sum(51:91)
```

8.5
```{r}
t <- seq(1:1000)
```
#a. How many data points from 8.1 to 8.4?
# 143 data points
```{r}
length1 <- length(a)
length2 <- length(l)
length3 <- length(i)
length4 <- length(k)
sum (length1+length2+length3+length4)
```
#b.Write the R code and its output from 8.1 to 8.4.

```{r}
a <- seq(1:100)
l <- seq(20,60)
i <- mean(20,60)
k <- sum(51:91)
```    

#c. For 8.5 find only maximum data points until 10.
```{r}
t <- 1:1000
answer <- max(t[t <- 10])
answer
```


#9.
```{r}
Filter(function(i) { all(i %% c(3,5,7) != 0) }, seq(100))
```

#10.
```{r}
rev(seq(1:100))
```

#11.
```{r}
upper_limit <- 25
num <- 1:(upper_limit - 1)
multiples <- num[num %% 3 == 0 | num %% 5 == 0]
multiples
```
#a. How many data points from 10 to 11?
# 11 data points
```{r}
length(multiples)
```
#b. Write the R code and its output from 10 and 11.
#

#12.
```{r}

