---
title: "Worksheet-4a in R"
author: "Necel Kate Noblezada"
date: "2024-10-30"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Using Loop Function
# for() loop
# 1. Using the for loop, create an R script that will display a 5x5 matrix as shown in Figure 1. It must contain vectorA = [1,2,3,4,5] and a 5 x 5 zero matrix.
# Hint Use abs() function to get the absolute value
```{r}
VectorA <- c(1, 2, 3, 4, 5)
matrix <- matrix(0, nrow = 5, ncol = 5)
for (i in 1:5) {
  matrix[i, ] <- VectorA
}

matrix
```

# 2. Print the string "*" using for() function. The output should be the same as shown in Figure
```{r}
r <- 5
for (i in 1:r) {
    cat(rep('"*"', i), "\n")
}
cat
```

# 3. Get an input from the user to print the Fibonacci sequence starting from the 1st input up to 500. Use repeat and break statements. Write the R Scripts and its output.
```{r}
start <- as.integer(readline(prompt = "Enter the starting number for the Fibonacci sequence: ")) 
a <- 0 
b <- 1 
repeat { 
  fib <- a + b 
  a <- b 
  b <- fib 
  if (fib >= start) { 
    cat(fib, " ") } 
  if (fib >= 500) 
    { 
    break }
}
```

# 4. Import the dataset as shown in Figure 1 you have created previously.
# a. What is the R script for importing an excel or a csv file? Display the first 6 rows of the dataset? Show your codes and its result
```{r}
library(readxl)
shoe_size <- read_excel("shoe_size.xlsx")
head(shoe_size)
```
# b. Create a subset for gender(female and male). How many observations are there in Male? How about in Female? Write the R scripts and its output.
```{r}
male_subset <- subset(shoe_size, Gender == "M")
female_subset <- subset(shoe_size, Gender == "F")

num_male <- nrow(male_subset)
num_female <- nrow(female_subset)

cat("Number of observations in Male:", num_male, "\n")
cat("Number of observations in Female:", num_female, "\n")
```
# c. Create a graph for the number of males and females for Household Data. Use plot(), chart type = barplot. Make sure to place title, legends, and colors. Write the R scripts and its result.
```{r}
gender_counts <- table(shoe_size$Gender)

barplot(gender_counts,
        main = "Number of Males and Females",
        xlab = "Gender",
        ylab = "Count",
        col = c("blue", "pink"),
        legend = rownames(gender_counts),
        beside = TRUE)
```

# 5. The monthly income of Dela Cruz family was spent on the following:
# a. Create a piechart that will include labels in percentage.Add some colors and title of the chart. Write the R scripts and show its output.
```{r}
expenses <- c(60, 10, 5, 25)
labels <- c("Food", "Electricity", "Savings", "Miscellaneous")
colors <- c("lightgreen", "lightblue", "lightyellow", "lightcoral")

pie(expenses,
    labels = paste(labels, "\n", round((expenses / sum(expenses)) * 100, 1), "%", sep = ""),
    col = colors,
    main = "Monthly Income Distribution of Dela Cruz Family")

```

# 6. Use the iris dataset.
# a. Check for the structure of the dataset using the str() function. Describe what you have seen in the output.
```{r}
data(iris)

str(iris)
```
# b.
```{r}
mean_values <- colMeans(iris[, 1:4])

print(mean_values)

# c.
species_counts <- table(iris$Species)

colors <- c("lightgreen", "lightblue", "lightcoral")

pie(species_counts,
    labels = paste(names(species_counts), "\n", species_counts, sep = ""),
    col = colors,
    main = "Species Distribution in Iris Dataset")
```
# d.
```{r}
setosa_subset <- iris[iris$Species == "setosa", ]
versicolor_subset <- iris[iris$Species == "versicolor", ]
virginica_subset <- iris[iris$Species == "virginica", ]

setosa_last_six <- tail(setosa_subset, 6)
versicolor_last_six <- tail(versicolor_subset, 6)
virginica_last_six <- tail(virginica_subset, 6)

print("Last six rows of Setosa:")
print(setosa_last_six)
print("Last six rows of Versicolor:")
print(versicolor_last_six)
print("Last six rows of Virginica:")
print(virginica_last_six)
```
# e.
```{r}
iris$Species <- as.factor(iris$Species)

plot(iris$Sepal.Length, iris$Sepal.Width,
     col = iris$Species,
     pch = 19,
     main = "Iris Dataset",
     sub = "Sepal Width and Length",
     xlab = "Sepal Length",
     ylab = "Sepal Width")

legend("topright", legend = levels(iris$Species), col = 1:3, pch = 19)
```








